# Explore 2 varable exercises:

In this problem set, you'll continue
to explore the diamonds data set.

```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
library(gridExtra)
data("diamonds")
```

Your first task is to create a
scatterplot of price vs x.
using the ggplot syntax.

```{r}
names(diamonds)
ggplot(aes(x = x, y = price), data = diamonds)+
  geom_point()+
  scale_x_continuous(limits = c(3,11))
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
  
```
Create a simple scatter plot of price vs depth.

```{r}
ggplot(aes(x = depth, y = price), data = diamonds)+
  geom_point(alpha = 1/100)+
  scale_x_continuous(breaks = seq(0,100,2))
cor.test(diamonds$depth, diamonds$price)
```

Create a scatterplot of price vs carat
and omit the top 1% of price and carat
values.

```{r}
ggplot(aes(x = carat, y = price), data = diamonds)+
  geom_point()+
  scale_x_continuous(limits = c(0,quantile(diamonds$carat, probs=.99)))+
  scale_y_continuous(limits = c(0,quantile(diamonds$price, probs=.99)))
?quantile
quantile(diamonds$carat, probs=.9)
```

Create a scatterplot of price vs. volume (x * y * z).
This is a very rough approximation for a diamond's volume.

Create a new variable for volume in the diamonds data frame.
This will be useful in a later exercise.

```{r}
diamonds$volume <- with(diamonds, x*y*z)

ggplot(aes(x = volume, y = price), data = diamonds)+
  geom_point()+
  scale_x_continuous(limits = c(0, 500))

diam_out <- subset(diamonds, diamonds$volume != 0 & diamonds$volume < 800)
cor.test(diam_out$volume, diam_out$price)
```
Subset the data to exclude diamonds with a volume
greater than or equal to 800. Also, exclude diamonds
with a volume of 0. Adjust the transparency of the
points and add a linear model to the plot. (See the
Instructor Notes or look up the documentation of
geom_smooth() for more details about smoothers.)

We encourage you to think about this next question and
to post your thoughts in the discussion section.

Do you think this would be a useful model to estimate
the price of diamonds? Why or why not?

```{r}
ggplot(aes(x = volume, y = price), data = diam_out)+
  geom_point(alpha = 1/20)+
  geom_smooth(method = 'lm', color = 'red')
?geom_smooth

```

Use the function dplyr package
to create a new data frame containing
info on diamonds by clarity.
Name the data frame diamondsByClarity

The data frame should contain the following
variables in this order.

   (1) mean_price
   (2) median_price
   (3) min_price
   (4) max_price
   (5) n

```{r}
gb_cl <- group_by(diamonds, clarity)
diamondsByClarity <- summarise(gb_cl,
                               mean_price = mean(price),
                               median_price = median(price),
                               min_price = min(price),
                               max_price = max(price),
                               n = n())

head(diamondsByClarity)
```

Your task is to write additional code to create two bar plots
on one output image using the grid.arrange() function from the package
gridExtra.


```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))
head(diamonds_mp_by_clarity)
diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
diamonds_mp_by_clarity$clarity <- factor(diamonds_mp_by_clarity$clarity, levels = c('I1' , 'SI1', 'SI2', 'VS1', 'VS2', 'VVS1', 'VVS2', 'IF'))
diamonds_mp_by_clarity <- arrange(diamonds_mp_by_clarity, clarity)

p1 <- ggplot(aes(x=clarity, y = mean_price), data = diamonds_mp_by_clarity)+
  geom_bar(stat = 'identity')+
  labs(title = 'mean price by clarity')

p2 <- ggplot(aes(x=color, y=mean_price), data = diamonds_mp_by_color)+
  geom_bar(stat='identity')+
  labs(title='mean price by color')
grid.arrange(p1, p2, ncol=2)
?diamonds
```





```{r}

```





