Lesson 3
========================================================

***

### What to Do First?
Notes: loads the file into a read.csv reader. It is a tab (\t) separated file

***

### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
pf <- read.csv("pseudo_facebook.tsv", sep = '\t')
names(pf)
dim(pf)

```

***

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
qplot(x = dob_day, data = pf) +
  scale_x_continuous(breaks=1:31) +
  facet_wrap(~dob_month, ncol = 3)

#facet_grid(vertical~horizontal) take variables to split on vertical and horizontal grid

#same plot:
#ggplot(aes(x = dob_day), data = pf) +
#  geom_histogram(binwidth = 1) +
#  scale_x_continuous(breaks = 1:31)

```


***

#### What are some things that you notice about this histogram?
Response:

***

### Moira's Investigation
Notes:

***

### Estimating Your Audience Size
Notes:

***

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

***

### Perceived Audience Size
Notes:

***
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

***

### Be Skeptical - Outliers and Anomalies
Notes:

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
#qplot(x = pf$friend_count, data = pf, xlim = c(0,1000))
#other possibility
 qplot(x = friend_count, data = subset(pf, !is.na(pf$gender)), binwidth = 25) +
  scale_x_continuous(limits = c(0,1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol = 1)

ggplot(aes(x = friend_count), data = subset(pf, !is.na(gender))) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}

```

### Exploring with Bin Width
Notes:

***

### Adjusting the Bin Width
Notes:

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
#see before
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
# see before
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

***

### Tenure
Notes:

```{r Tenure}
qplot(x = tenure, data = pf, color = I('black'), fill = I('#099DD9'), binwidth = 30)

#tenure in years:
qplot(x = tenure/365, data = pf, color = I('black'), fill = I('#099DD9'), binwidth = .25,
      xlab = "tenure (years)",
      ylab = "frequency")+
  scale_x_continuous(breaks = seq(0,10,1), limits = c(0,10))

```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}

```

***

### Labeling Plots
Notes:

```{r Labeling Plots}

```

***

### User Ages
Notes:

```{r User Ages}
qplot(x = age, data = pf, color = I('black'), fill = I('#099DD9'), binwidth = 2)+
  scale_x_continuous(breaks = seq(0,120,10))
  
```

#### What do you notice?
Response:

***

### The Spread of Memes
Notes:

***

### Lada's Money Bag Meme
Notes:

***

### Transforming Data
Notes:

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#install.packages("gridExtra")
library(gridExtra)
p1 <- qplot(x = friend_count, data = pf, color = I('black'), fill = I('#099DD9'),
           binwidth = 10)+
  scale_x_continuous(breaks = seq(0,1000,10), limits = c(-1,1000))
p2 <- qplot(x = friend_count, data = pf, color = I('black'), fill = I('#099DD9'),
           binwidth = 10)+
  scale_x_continuous(breaks = seq(0,1000,10), limits = c(0,1000))+
  scale_y_log10(name = 'log')

p3 <- qplot(x = friend_count, data = pf, color = I('black'), fill = I('#099DD9'),
           binwidth = 10)+
  scale_x_continuous(breaks = seq(0,1000,10), limits = c(0,1000))+
  scale_y_sqrt(name = 'sqrt')

p4 <- qplot(x = log10(friend_count + 1), data = pf)

grid.arrange(p1, p2, p3, p4, ncol = 1)

# transforming X data with ggpplot:

p1 <- ggplot(aes(x = friend_count), data = pf)+geom_histogram() 
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Percentage of users with that friend count')

# same with qplot:
qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), geom='freqpoly',
      color = gender, binwidth = 10)+
    scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Percentage of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x = www_likes, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), geom='freqpoly',
      color = gender)+
    scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('likes on the webt') +
  ylab('Percentage of users')+
  scale_x_log10()

#in python: use the seaborn library with parameter "hue" to do the same as color = gender
# à écrire dans mon document

by(pf$www_likes, pf$gender, sum)


```


***

### Box Plots
Notes:

```{r Box Plots}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')+
  scale_y_continuous(limits = c(0, 1000))

# better way, because by using limits we change the data used to 
# calculate the quartile limits

qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')+
  coord_cartesian(ylim = c(0, 1000))

by(pf$friend_count, pf$gender, summary)
names(pf)
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot')+
  coord_cartesian(ylim = c(0, 200))

by(pf$friendships_initiated, pf$gender, summary)
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}

```

Response:

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes >0) # make a summary of the function used

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

63947/length(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

Response:

***

### Analyzing One Variable
Reflection:
```{R Exercise on diamond dataset}

library(ggplot2)
#Exo 1:
data("diamonds")
summary(diamonds)
dim(diamonds)
str(diamonds)
levels(factor(diamonds$color))
?diamonds

# Exo 2:
qplot(x = price, data = diamonds, binwidth = 500)
summary(diamonds$price)
length(subset(diamonds$price, diamonds$price >= 15000))

qplot(x = price, data = diamonds, binwidth = 5,
      color = I('black'), fill = I('#099DD9'))+
  scale_x_continuous(breaks = seq(0, 2000, 25), limits = c(400,1000))

# break price histogram by cut:
str(diamonds$cut)
qplot(x = price, data = diamonds)+
  facet_wrap(~cut, ncol = 3)

by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$cut, max)

# free scale for facet_wrap:
?facet_wrap

qplot(x = price, data = diamonds)+
  facet_wrap(~cut, ncol = 3, scales = 'free_y')


# price/carat faceted by cut:

qplot(x = (price/carat), data = diamonds, binwidth = .03)+
    scale_x_log10(limits = c(1100, 11000))+
  facet_wrap(~cut, scales = "free_y", ncol = 2)

# boxplot of price by clarity, i would use ggplot instead of qplot ;)

ggplot(aes(x = clarity, y = price), data = diamonds)+
  geom_boxplot()

ggplot(aes(x = clarity, y = price), data = diamonds)+
  geom_boxplot()+
  scale_y_log10()

by(diamonds$price, diamonds$color, summary)
IQR(diamonds$price[diamonds$color == "D"])

by(diamonds$price, diamonds$color, IQR)
?diamonds

ggplot(aes(x = color, y = price/carat), data = diamonds)+
  geom_boxplot()+
  scale_y_log10(limits = c(1000, 10000), breaks = seq(1000, 10000, 1000))

# frequency polygon of carat:

ggplot(aes(x = carat), data = diamonds)+
  geom_freqpoly(binwidth = 0.01)+
  scale_x_continuous(limits = c(1.9, 2.3), breaks = seq(0, 2, .1))

install.packages("tidyr")
install.packages("dplyr")

```

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!