# Exploratory data analysis in the Red wine composition Dataset

## Introduction to the dataset:

The Dataset chosen for this project is the red wine composition dataset.
I chose this dataset because I am really interrested in wine and i thought it
would be really interresting dataset. This dataset comes from 
[Cortez et al. 2009 [1]](#Cortez). It present different caracteristics
of red and white wine. Only the Red wine data was explored in this document.
This dataset contain 1599 wine analyses. Each sample was analyzed for different 
[characteristics](#characteristics) and then the wine quality was assessed by at
least 3 expert giving a grade between 0 (very bad) and 10 (very excellent). 
More informations can be found in the [M&M](./M&M) file.

### Wine characteristics studied: <a name="characteristics"></a>

The variables present in this dataset are:
   
   1. fixed acidity (tartaric acid - g / dm³)
   2. volatile acidity (acetic acid - g / dm³)
   3. citric acid (g / dm³)
   4. residual sugar (g / dm³)
   5. chlorides (sodium chloride - g / dm³
   6. free sulfur dioxide (mg / dm³)
   7. total sulfur dioxide (mg / dm³)
   8. density (g / cm³)
   9. pH
   10. sulphates (potassium sulphate - g / dm3)
   11. alcohol (% by volume)
   Output variable (based on sensory data): 
   12 - quality (score between 0 and 10)


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(gridExtra)
library(GGally)
library(dplyr)
library(memisc)
library(scales)
# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red <- read.csv('wineQualityReds.csv')

```


## Univariate Plots Section

Lets look at the dataset before plotting anything. Looking at the structure of
the dataset gives a good sense of it.

```{r echo=FALSE}
str(red)
```

One thing I first notice is that we have 13 variables instead of 12. The
unwanted variable is the first one: X. The summary of this variable has been 
computed below: 

```{r echo=FALSE}
summary(red$X)
```

In this summary we see that the variable range from 1 to 1599 which is the 
number of samples analysed in this dataset. To check if this column is just an
artefact from a row numbering in our csv file that has been misinterpreted 
during importation we can compute an histogram of this variable with a binwidth 
of 1:

```{r echo=FALSE}
ggplot(aes(x = X), data = red)+
  geom_histogram(binwidth = 1)
```

We clearly see that each X value has a frequency of 1. We can conclude that this
column can be removed from the dataset as it only numbered the sample and 
doesn't represent anything.

```{r echo=FALSE}
red <- subset(red, select = -X)
```

Another "problem" could be the mapping of quality as an integer and not a 
factor. Quality is a number representing a qualitative variable. The use of
integer type on this variable might be interresting when modeling or doing 
scatterplot, but it is a bit misleading. I added a column to the dataset named 
quality.factor to code the quality variable as a factor while keeping the 
integer mapping in case it is needed later.

```{r echo=FALSE}
red$quality.factor <- factor(red$quality, levels = c(0,1,2,3,4,5,6,7,8,9,10))
```

Next let's compute a summary of the dataset:

```{r echo=FALSE}
summary(red)

```

Looking at the repartition of value of the different variables, we can see that 
 fixed.acidity, residual.sugar, chlorides, free.sulfur.dioxide, 
 total.sulfur.dioxide and sulphates shows some really high maximal value 
 compared to the 3rd quartile value. These outlier could be problematic later.

We will now look at the shape of each variable. First I will look at non factor 
variable. I find Boxplots to be a good way to visualise distribution feature 
such as quartile and outliers, but they aren't as detailled as histograms.
I will make these two type of plot side by side. To avoid repetition I created a function named `distrib` that make the plot.

```{r}
# function to display a boxplot and an histogram side by side.
# input:
#       - df: dataframe
#       - x1: variable to plot
#       - bi: number of bins to use for the histogram 
distrib <- function(df = red, x1, bi = 30) {
  p1 <- ggplot(data = df, aes_string(x = x1)) +
    geom_histogram(bins = bi)
  p2 <- ggplot(data = df, aes_string(y = x1))+
    geom_boxplot(aes(x = 1))+
    stat_summary(aes(x = 1), fun.y = "mean",
                 geom = "point", color = "red", shape = 3, size = 4)
  grid.arrange(p1, p2, ncol = 2)
}
```


```{r}
distrib(x1 = "fixed.acidity")

```

fixed acidity show a skewed distribution with a long tail between 12 and 16.
Most of the sample have a fixed acidity value between 6 and 10 g/dm³

```{r}
distrib(x1 = "volatile.acidity")
```

Volatile acidity also show a long tailled distribution, but with a lot less 
outliers. Most of the sample have a value of 0.2 to 0.8 g/dm³.

```{r}
distrib(x1 = "citric.acid")

```

The citric acid distribution is again skewed, but only shows one outlier. The 
majotiry of the sample have a concentration of 0 to 0.65 g/dm³. The histogram
show that there is lots of wine with a concentration near 0 g/dm³, then the 
number of sample with a concentration between 0 and 0.25 g/dm³ decrease until 
0.25 g/dm³ around the median, we have an increase in the number of sample with 
such a concentration. There is another high count for sample with citric acid concentration around 0.5g/dm³. It is interresting to see these high count. There
might be a biological reason.

```{r}
distrib(x1 = "residual.sugar")
```

Residual sugar shows a highly skewed distribution with lots of ponctual extreme values. Most of the variation appear below 4 g/dm³. I will zoom on the part 
between 1 and 4 g/dm³

```{r}
ggplot(data = red, aes(x = residual.sugar))+
  geom_histogram()+
  xlim(1, 4)
```

Zooming allow us to see that the distribution of resdual sugar is bell shaped 
between 1.25 and 3 g/dm³. After 3 g/dm³ the number of sample decrese slowly.


```{r}
distrib(x1 = "chlorides")
```


The chloride distribution is even more extremly skeawed than residual sugar. I 
will zoom on the part with chlorides concentration between (0.025 , 0.13) g/dm³.

```{r}
ggplot(data = red, aes(x = chlorides))+
  geom_histogram()+
  xlim(0.025, 0.13)
```

In this range, the data repartition looks gaussian.

```{r}
distrib(x1 = "free.sulfur.dioxide")
```

The free sulfur dioxide distribution also sow some outliers, but is much less 
skewed than the example before.

```{r}
distrib(x1 = "total.sulfur.dioxide")
```

The total sulfur dioxide repartition ooks similar to the free sulfur dioxide repartition. But there is 2 outliers with high SO2 concentrations.
When we zoom on the part with most of the frequency:

```{r}
distrib(x1 = "density")
```

density shows a gaussian type density curve. Most of the sample have a density
of (0.993, 1.000) g/dm³.

```{r}
distrib(x1 = "pH")
```

The pH also show a gaussian density curve centered in 3.310.

```{r}
distrib(x1 = "sulphates")
```

The sulphate distribution show some outliers, most of the values are found below
1-1.20 g/dm³.

```{r}
distrib(x1 = "alcohol")
```

alcohol show a skewed distribution.

 Looking at thoses variables, we have seen variables with very little variation 
between the 1st and 3rd quaritle: chlorides, residual sugar and sulphates.
The density and pH parameter have a centered distribution while other parameters
have in general a distribution skewed toward large value.

The next step is to look at the repartition of wine between the different 
quality, an histogram is the best way to have an overview.

```{r echo=FALSE}
ggplot(aes(x = quality.factor, y = ..count../sum(count)), data = red)+
  geom_bar(stat = "count")
```

In this dataset we have wine with quality ranging from 3 to 8. No wine have 
really bad quality (0 to 2) and no one have really good quality (9, 10). The 
majority of wine fall into the medium quality, ~80% of the wine were of quality 
5 and 6. The quality distribution look balanced in its range, but have more wine
with a quality score higher than 6, than quality lower than 5.

## Univariate Analysis:

### What is the structure of your dataset?

This dataset contain the composition analysis and quality score of 1599
wines. The composition parameters found in this dataset are:

 - fixed.acidity
 - volatile.acidity
 - citric.acid
 - residual.sugar
 - chlorides
 - free.sulfur.dioxide
 - total.sulfur.dioxide
 - density
 - pH
 - sulphates
 - alcohol

The quality has been recorded as a grade from 0 (very bad) to 10 (very good).
Looking at the repartition of wine quality, we have seen that scores lies 
between 3 and 8 with the majority of wine tested having a medium quality (5-6), 
and that there is more wine with quality 7 - 8 than quality 3 - 4.

### What features of is there in your dataset?

One interresting feature is obviously the quality of the wine as it is the main
reason one would drink wine. But looking at the repartition of the different 
variables some showed a really low variability and it might be difficult to 
easily find a relationship describing quality.
Another interresting part could be to look at the relations between variables.
For instance, we have 4 variables related to acidity of the wine:

 - fixed.acidity
 - volatile.acidity
 - citric.acid
 - pH
 
 We might be able to find a relation between them. Also it could be interresting
 to look at relations  between free.sulfur.dioxide, total.sulfur.dioxide and 
 sulphates as they are related to SO2 concentration.

### Did you create any new variables from existing variables in the dataset?

I just created a factorial version of the quality variable. In other parts of 
the analysis I might create a new variable accounting for quality with only 4 
level:<a name="category"></a>

- bad: qualtity 3 and 4
- medium low: quality 5 
- medium high: quality 6
- good: quality 7 and 8

Because wine with low and high quality are not a big part of our sample, it 
could make some visualisation easier to read. This idea is not fixed yet we will
see if it shows any interresting relations compared to the actual quality 
system.

## Bivariate Plots Section

First lets use the ggcorr function to go a bit faster in representing the 
variables against each other (variables were plotted against each other, but 
that project will look at linear correlation as a guide to analyse relations
between variables):

<a name="correl">
```{r echo=FALSE, Bivariate_Plots, fig.height=8, fig.width=8}

ggcorr(red[,1:11], digits =  3, label = TRUE, label_round = 3,
       layout.exp = 0.05 , nudge_x = -0.5)
```
</a>

Looking at the linear correlation  coefficient gives us a good idea of linear 
relationships. I will first talk about relationships that have a linear 
correlation around |0.5| or better, and then relationship between |0.2| and 
|0.4|.

The strongest linear correlations are:

- **total.sulfur.dioxyde** and **free.sulfur.dioxyde** with a correletion of 
    **0.668** this correlation seems logical as free.sulfur.dioxyde represent a 
    part of total.sulfur.dioxyde.

<a name="positive"></a>

- **citric.acid** with **fixed.acidity** with a correletion of **0.672**
- **pH** with **fixed.acidity** with a correletion of **-0.683**
- **citric.acid** with **pH** with a correletion of **-0.542**
- **citric.acid** with **volatile.acidity** with a correlation of **-0.552**

These four correlation are logical as they happen between different acids or 
between acids and the pH which represent the acidity of the wine. The negative 
correletion coeficient between pH and other acids comes from the fact that the 
lower the pH the more acid the solution is.

- **density** with **fixed.acidity** with a correletion of **0.668**
- **density** with **alcohol** with a correletion of **-0.496**

We know that  the desity is know to be related to alcohol and the sugar content 
(see [M&M](./M&M) file), it is therefore not surprising to find a relation 
between density and alcohol. Despite this relation, the linear relationship 
between density and residual.sugar is not as strong (corr coef of 0.355) as I 
would expect it to be. If we plot density against alcohol and residual sugar to
see the difference.

```{r}
pa <- ggplot(data = red, aes(x = alcohol, y = density))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)

pz <- ggplot(data = red, aes(x = residual.sugar, y = density))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)

grid.arrange(pa, pz, nrow = 1)

```

We already discussed the high number of outliers in the residual sugar variable.
On the scaterplot we can see that most of the samples have a residual sugar 
below the 4 g/dm³.We will remove sample with residual sugar higher than 4 to see
if we can have a better correlation. The 4 g/dm³ is an arbitrary value decided 
after having a look at the distribution of residual sugar, it doesn't come from a
ny statistical method of outlier detection.

```{r}
lower_rs <- subset(red, residual.sugar <= 4)
co <- cor(lower_rs$density, lower_rs$residual.sugar)
co1 <- sprintf("r = %.3f", round(co, digits = 3))
ggplot(data = lower_rs, aes(x = residual.sugar, y = density))+
  geom_point()+
  geom_smooth(method = "lm")+
  annotate("text", x = 1.25, y = 1.001, label = co1)

```

By removing the data point for sample with residual sugar over 4 we can see that
the correletion coefficient has been slightly improved, but we only improve from
r = 0.355 to r = 0.391. We see that values are only found for some specific 
residual sugar value. It probably comes from the measurment method and the 
precision of the equipment used that probably round the residual sugar value to 
the hundredth of gram/dm³.

An interresting relationship is the density and fixed.acidity relationship.

```{r}
ggplot(data = red, aes(x = fixed.acidity, y= density))+
  geom_point()+
  geom_smooth(method = 'lm', se = FALSE)
```

This relationship is interresting as fixed.acidity is in majority formed by the
tartric acid which comes from the fruit directly ([tartric acid in wine](https://en.wikipedia.org/wiki/Acids_in_wine#Tartaric_acid)) and not the 
fermantation, so part of the wine density is a direct result of the grape 
quality at harvest.

Some of the less pronounced linear relationships tendencies can be found 
between:

- **fixed.acidity** and **volatile acidity**, coeficient of **-0.256**
- **pH** and **volatile.acidity**, coefficient of **0.235**
- **volatile.acidity** and **sulphates**, coefficient of **-0.261**
- **volatile.acidity** and **alcohol**, coefficient of **-0.202**
- **citric.acid** and **chlorides**, coefficient of **0.204**
- **citric.acid** and **density**, coefficient of **0.313**
- **citric.acid** and **sulphates**, coefficient of **0.313**
- **pH** and **alcohol**, coefficient of **0.206**

These coeficients show the tendency of a linear relation between componant of 
acidity and other characteristics observed in the wine. Other coefficient 
greater than |0.2| can be found, but as they don't relate to the acidity 
charachteristics I decided To let them aside for this report.

Next we should look at the quality variable and its relation to other 
characteristics. I will first look at relation with quality to see interresting
relationships. Once I find interresting relationships, I will explore the 
boxplot and compare boxplot with 10 quality classes and the boxplot with the 4 
quality classes discussed [before](#category).

```{r echo=FALSE, fig.height=12}
red$quality.four <- factor(ifelse(red$quality < 5, "low",
                           ifelse(red$quality > 6, "high", 
                           ifelse(red$quality == 5, "medium low",
                                  "medium high"))))
red$quality.four <- factor(red$quality.four , 
                      levels = c("low", "medium low", "medium high", "high"))


plo <- list()
l <- 1
for (i in names(red[, c(-12, -13, -14)]))
{
  plo[[l]] <- ggplot(aes_string(x = "quality.factor", y = i ), data = red)+
    geom_boxplot()
  l <- l + 1
}
do.call(grid.arrange, c(plo, ncol = 2))
```

By looking at thoses boxplot, interresting trends are seen for volatile.acidity, 
citric.acid, alcohol. Other variables show lesser impressive pattern or 
trend. 

First, let's look at these boxplot, we will also compare them to a 
boxplot using the 4 class category for quality discussed [before](#category).

```{r}
two_box <- function(df = red, x1 = "quality.factor", 
                    x2 = "quality.four", y) {
  p1 <- ggplot(data = df, aes_string(x = x1, y = y))+
    geom_boxplot()+
    stat_summary(fun.y = "mean",
                 geom = "point", color = "red", shape = 3, size = 2)
  p2 <- ggplot(data = df, aes_string(x = x2, y = y))+
    geom_boxplot()+
    stat_summary(fun.y = "mean",
                 geom = "point", color = "red", shape = 3, size = 2)
  grid.arrange(p1, p2, nrow = 1)
                    }

two_box(y = "volatile.acidity")

```


Volatile acidity show a decrease in acetic acid with wine having a better 
quality. According to the wikipedia [article](https://en.wikipedia.org/wiki/Acids_in_wine#Other_acids) about acid in
wine, acetic acid is produce if the wine as been exposed to oxygen during its 
fermentation. This acid is associated to vinegar taste which is equivalent to a 
bad quality and not an enjoyable drink. This article also state that amount over 
600mg/l (=0.6g/dm³) are easily detected by most people. It then seems logical 
that wine expert can qualified wine with acetic acid up to 0.4g/dm³ as they are 
trained to detect imperfection in wine. We see that a wine of medium quality 
show in median 0.6g/dm³. Interrestingly, it seems that acetic acid concentraton
below 0.4g/dm³ aren't making any differences as the median and repartition of 
acetic acid concentration between wine of quality 7 and 8 are similar.
volatile acidity also show outliers with high volatile acidity while they have a 
quality of 8, maybe these wine with high volatile acidity have some other 
charachteristics that could mask  or compensate for it. The boxplot using the 4 
quality level only show the decreasing trend. It hide the plateau in 
concentration between quality 7 and 8.

```{r}
two_box(y = "citric.acid")
```


Citric acid shows a positive relatioship, wine with higher concentration seems 
to be rated with a higher quality. Citric acid is reported adding freshness 
taste to wine. We see that for quality over 7 the citric acid doesn't seems to 
help increase the quality score. The 4 category boxplot doesn't add much, it is
interresting to see the difference between mean and median in low (3-4) quality.
Because of the highly skewed distribution of concentration, the mean doesn't 
show as much differences as the median. 

```{r}
two_box(y = "alcohol")
```

<a name="alc"></a>
The alcohol relationship to quality is really interresting because wine with 
quality between 3 and 5 have a similar median of alcohol concentration, but 
wine with quality over 5 it appears that higher alcohol concentration is 
correlated with higher quality score. Again, using only 4 quality level doesn't 
add to the interpretation, but it could help modelise relations because 
category with low sample number would have more sample and thus could help have 
stronger statistical meaning.

Another interresting fact is the range and number of outlier found for a lot of 
variables. For example residual sugar:

```{r}
p1 <- ggplot(data = red, aes(x = quality.factor, y = residual.sugar))+
  geom_boxplot()+
  stat_summary(fun.y = "mean",
               geom = "point", shape = 3, color = "red")+
  ggtitle("with outliers")

p2 <- ggplot(data = subset(red, red$residual.sugar <= 4), aes(x = quality.factor, y = residual.sugar))+
  geom_boxplot()+
  stat_summary(fun.y = "mean",
               geom = "point", shape = 3, color = "red")+
  ggtitle("without residual sugar > 4g/dm³")

grid.arrange(p1, p2, nrow = 1)
```

We see that by removing extreme value, we have a better view of the repartition 
of the residual sugar variable, but in the case of residual sugar it doesn't 
help us find any trend.
The fact that residual sugar show so many outliers could be explained because 
there is different style of wine and in each category quality is defined a bit 
differently. [Wine tasting](https://en.wikipedia.org/wiki/Wine#Tasting) is a 
complicated task and quality of a wine depend on the different general quality, 
but is also dependend on a wide array of chemicals changing its taste.

By reducing the number of category for the quality variable we aren't showing 
trend that where hidden, but we have group containing more sample, this can give
a better statistical weight to each category feature.


## Multivariate Plots Section

To start with multivariate plotting, I first want to look at relation between 
pH and citric acid because of its corelation coefficient, and color the point by 
the volatile acidity.

```{r echo=FALSE}
ggplot(data = red, aes(x = pH, y = citric.acid, color = volatile.acidity))+
  geom_point()+
  scale_colour_gradientn(colours = rainbow(3))
```

This first graph show a decreasing relationship between citric acid and pH, but 
also that for the same pH the more citric acid, the less volatile acidity.
It seems logical as pH comes from the acidity of every acid found in the wine.

Next I want to see the combination of volatile acidity and citric acid over the 
different quality of wine.

```{r echo=FALSE}

ggplot(data = red, aes(x = citric.acid, y = volatile.acidity, 
                             color = quality.factor))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

In this graph we see the volatile acidity and the citric acid, looking at the 
color gradient and the regression lines, we can see that high quality wine have 
a lower volatile acidity content, we also see that volatile acidity and citric
acid share a decreasing relationship, the more citric acid, the less volatile 
acidity in the wine.

In the next graph I will look at the relation found in the first graph, but 
facet it by quality score.

```{r echo=FALSE}
ggplot(data = red, aes(x = pH, y = citric.acid, color = volatile.acidity))+
  geom_point()+
  facet_wrap(~quality)+
  scale_colour_gradientn(colours = rainbow(3))

```

In this graph we see that within each quality score the whole range of volatile
acidity and citric acid can be found. Another interresting remark is the 
differences in number between the middle category (quality = 5-6) and the other 
category. Another interesting fact in this dataset is that category with lower 
wine number have a gap between wine with citric acid concentration lower than 
0.125g/dm³ and higher than 0.20-0.25g/dm³. Let's try to look at the same graph
but faceted following the aggregation of lower quality and higher quality wine.

```{r echo=FALSE}
ggplot(data = red, aes(x = pH, y = citric.acid, color = volatile.acidity))+
  geom_point()+
  facet_wrap(~quality.four)+
  scale_colour_gradientn(colours = rainbow(3))
```

With these category having higher number of wine we see the separation into 
groups more easily. In the high and low quality groups these two groups are 
separate. I have no interpretation of this observation.

Next lets look at the fixed acidity, pH and citric acid at the same time.

```{r}
ggplot(data = red, aes(x = pH, y = fixed.acidity, color = citric.acid))+
  geom_point(position = position_jitter())+
  facet_wrap(~quality)+
  scale_colour_gradientn(colours = rainbow(3))
```

We can see the positive correlation between fixed.acidity and citric acid found 
[before](#positive). Fixed acidity is the measure of Tartric acid. This acid is 
considered one of the [most important in wine](https://en.wikipedia.org/wiki/Acids_in_wine#Tartaric_acid). Looking at 
the repartition of tartric acid within the pH range, it is definitely an 
important factor in acidity of the wine.

As reported in the wikipedia article, tartric acid and citric acid comes from 
the grapes composition, it could be interresting to have information on the 
variety used to produce the wine tested. The [soil composition](https://en.wikipedia.org/wiki/Acids_in_wine#Tartaric_acid) also has
an impact on grape composition, the soil informations could also give 
interresting insight on the relation between wine acids.

Next I will try to see the repartition of residual sugar and alcohol by 
category. Residual sugar being related to the alcohol fermentation and the 
inithial sugar content we could see interresting trend, even if no linear trend 
should be found (see [correlations](#correl)).

```{r echo=FALSE}
ggplot(data = red, aes(x = alcohol, y = residual.sugar, color = quality.factor))+
  geom_point(position = position_jitter())+
  scale_y_log10()+
    facet_wrap(~quality.four)
```

I chose to plot the four quality category, and display the original quality 
category with color. In these visualisation, we see that residual sugar and 
alcohol doesn't have any relation.

If we look at the relation between alcohol and quality, we can see that the 
range of alcohol is wider for wine wit high and medium high quality.

Next We could try to see if linear regression could provide a way to modelise 
the quality of a wine using the different variable. After seeing the
relations between quality and the different variable not showing clear trend or
showing non linear tendency we shouldn't expect too much, but it could be 
interresting to see such a model, also we could compare the full model with 
model using only alcohol, and model using some or all the acid related variable.

```{r echo=FALSE}

m1 <- lm(quality ~ volatile.acidity, red)
m2 <- update(m1, ~ . + citric.acid)
m3 <- update(m2, ~ . + fixed.acidity)
m4 <- update(m3, ~ . + pH)
m5 <- lm(quality ~ alcohol, red)
m6 <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid 
           + residual.sugar + chlorides + free.sulfur.dioxide 
           + total.sulfur.dioxide + density + pH + sulphates + alcohol, red)

mtable(m1, m2, m3, m4, m5, m6, sdigits = 3)
```

The first remark is that all these models show low R squared (0.153 to 3.61). So
these models are not really good, the variables account for less than 1/2 of the 
variance in quality.
I will next compare the model with only acid related variables (models m1-m4),
then look at the last two models.

Looking at the acids models, we observe no difference in R^2 between model 
using volatile acidity, fixed acidity and citric acid. 
We also see that in the first three models volatile acidity is the only 
significant variable. In the third model pH add a small improvement
in R^2 but it is really minimal. This can be explained by looking at the graph 
showed before and that acid parameters show some relations. These variables 
aren't showing linear trend with quality and they seem to be higly related to 
each other. This lack of independence makes the added information coming from 
one of these variable inferable from the result of one of the other variables.

We already [saw](#alc) that alcohol showed an interresting trend with quality, 
the model m5 show a R^2 higher than the one from the alcohol data. In this model 
alcohol account for 22% of the variance in quality.

In the complete model we see that the created model account for only 36% of the 
variance in quality. In this model, the variable showing the most significance 
are:

- volatile acidity
- alcohol
- chlorides
- total.sulfur.dioxide
- sulphates


## Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
ggplot(data = red, aes(x = quality.factor, y = alcohol))+
  geom_jitter(color = "#008080", alpha = 0.4)+
  geom_boxplot(color = "#FF4500", alpha = 0.5)+
  stat_summary(fun.y = "mean",
               geom = "point", shape = 8, color = "#800000", size = 4)+
  labs(x = "quality", y = "alcohol (% by volume)",
       title = "boxplot of the alcohol concentration \n between wine quality")+
  theme(plot.title = element_text(face = "bold", size = 15))
```

### Description One

This boxplot show the interresting relation between alcohol and quality. The 
positive relationship between alcohol and quality for wine with quality over 5
is the reason it is a significant factor in the linear regression we did before.
I added the sample in the background to show the repartition of sample within 
category. We observe the high number of sample falling into the 5-6 quality 
category. Category 5-6 account for 80% of the data points.

### Plot Two
```{r echo=FALSE, Plot_Two}

ggplot(data = red, aes(x = citric.acid, y = volatile.acidity, 
                             color = quality.factor))+
  geom_point(alpha = 0.8)+
  geom_smooth(method = "lm", se = FALSE)+
    labs(x = "citric acid (g/dm³)", 
       y = "volatile acidity (acetic acid - g / dm³)",
       title = "volatile acidity over citric acid concentration
per wine quality", colour = "quality")+
  theme(plot.title = element_text(face = "bold", size = 15))

```

### Description Two

This scatter plot present the volatile acidity and the citric acid per quality.
We see the negative relatioship between volatile acidity and citric acid 
found with the negative correlation coefficient (r = -0.552) to be true for 
every category, each regression line being decreasing. It means that for higher 
volatile acidity level, we have lower citric acid concentration. Looking at the
quality regression line, and point color, we observe that for lower quality 
wine tend to contain more volatile acidity.

The strong linear correletion coefficient show that citric acid and colatile 
acidity are not independent variables. This dependent relationship between acid 
are part of the explaination for having only two of these variable significant
in our linear modelisation.

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(data = red, aes(x = pH, y = fixed.acidity, color = citric.acid))+
  geom_point(position = position_jitter())+
  facet_wrap(~quality)+
  geom_smooth(method = "lm", se = FALSE)+
  scale_colour_gradientn(colours = rainbow(3))+
  labs(x = "pH", y = "fixed acidity (tartric acid - g / dm³)",
       title = "fixed acidity distribution within pH 
colored by citric acid concentration 
and faceted by wine quality", colour = "citric acid\n (g/dm³)")+
  theme(plot.title = element_text(face = "bold", size = 15))
```

### Description Three

We see again the relation between some of the acid variables. We can see the 
positive relationship between citric acid and tartric acid (= fixed acidity)
(r = 0.672), and their relation to pH.
fixed acidity has a decreasing relationship to ph (r = -0.683). 
pH also have a decreasing relationship with citric acid (r = -0.542). These 
relationship are decreasing because pH scale is lower when the acidity 
increases.

Looking at the different quality facet, we see similar pattern. 

## Reflection

I looked at the composition of wine and the repartition of chemical composition
within different wine. We also compared the quality of wine and tried to do 
a simple linear modelisation of quality using the different variables.

An important part of the investigation relied on the relations of different 
variables related to the acidic composition of wine. We have found that in the 
sample of wine tested the different acid can be related. These relation between 
acids in wine made the use of more than 2 of these variables not significant to
modelise quality. To continue the exploration, it could be interresting to 
modelise one of the acid variable using the other acid related variable as 
input.

The linear modelisation of the quality wasn't of good quality. The relation 
between some of the variables and the non linear trend showed between quality 
and other variable didn't help to find a good fit. Also the use of quality,
a qualitative variable as a dummy output variable might also not be the best 
modelisation technic. We could try to use different clustering or modelisation
technics. Also we probably need to reduce the number of dimension of this 
dataset to take into account the relation between variables found. Maybe some 
additionnal variables such as the soil type, soil composition or other 
compositions informations could help modelise more precisely the quality of 
wine.
Another information is that the dataset is strongly unbalanced between the 
quality categories and that 80% of the wine falls into the medium quality 
categories. We have only a few wine sample for bad and good quality. The 
unbalanced characteristics of this dataset are also part of the difficulty to 
accurately modelise quality. With such dataset, the model could be overtrained
to find medium quality wine.

## Bibliography:

<a name="Cortez" >[1] :   P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. 
 Reis. Modeling wine preferences by data mining from physicochemical properties.
 In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.</a>
  
[2] : https://en.wikipedia.org/wiki/Acids_in_wine